---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# didwrappers

<!-- badges: start -->
<!-- badges: end -->

The goal of didwrappers is to extend the functionality of the `did` package, specifically for within treatment-group heterogeneity and unbalanced data. 

## Installation

You can install the development version of didwrappers:

``` {r}
# install.packages("devtools")
devtools::install_github("ransiw/didwrappers")
```

## Example

This is a basic example generating code with a sample dataset:

```{r example}
library(didwrappers)
```

Simulate sample data

```{r}
simdata = sim_data()
head(simdata[,c("unit","time","treatg","dosage","y")])
```

Run an att_it() object and tabulate it
```{r}
attobject = att_it(yname = "y", tname = "time", gname = "treatg", idname ="unit", data = simdata)
attdf = attit_table(attobject)
head(attdf)
```

We can now aggregate all post-treatment effects to the unit level. 

```{r}
agtobject = aggite(attobject,type="unit")
aggite_table(agtobject)
```

Now sample a data frame with different dosage amounts and aggregate to the unit level.

```{r}
attobject = att_it(yname = "y", tname = "time", gname = "treatg", idname ="unit", cohortnames = "dosage", data = sim_data(dosage = rep(c(1,2),each=5)))
agtobject = aggite(attobject,type="unit")
aggite_table(agtobject)
```

We could also aggregate to the dosage level by specifying dosage as the `type`. 

```{r}
agtobject = aggite(attobject,type="dosage")
aggite_table(agtobject)
```

The functions can also accommodate unbalancedness. We can impose unbalancedness in simulated data. For example we could drop the first unit before time period 11, which essentially eliminates the unit.

```{r, message=FALSE}
simdata = sim_data()
simdata = simdata[!(simdata[,"unit"]==1 & simdata[,"time"]<11),]
attobject = att_it(yname = "y", tname = "time", gname = "treatg", idname ="unit", panel = FALSE, data = simdata)
```


```{r, message=FALSE}
agtobject = aggite(attobject,type="unit", na.rm = TRUE)
aggite_table(agtobject)
```

We could also consider the effect of truncation of the outcome variable at 100. Truncation will affect units whose baselevels are high.

```{r}
simdata = sim_data()
simdata$y[simdata$y > 100] <- 100
unique(simdata[simdata$baselevel>75,c("unit","treatg", "baselevel")])
```


We now calculate the unit level treatment effect for these.

```{r}
attobject = att_it(yname = "y", tname = "time", gname = "treatg", idname ="unit", panel = FALSE, data = simdata)
agtobject = aggite(attobject,type="unit", na.rm = TRUE)
aggite_table(agtobject)
```

Units 9 and 10 have negative estimates caused by the truncation. This could be alleviated by matching units at the baselevels, which can be done with matching units at the pre-treatment levels. 

```{r}
attobject = att_it(yname = "y", tname = "time", gname = "treatg", idname ="unit", panel = FALSE, xformla = ~y, data = simdata)
agtobject = aggite(attobject,type="unit", na.rm = TRUE)
aggite_table(agtobject)
```

